An FMI-Based Initialization Plugin for INTO-CPS Maestro 2\thanks{We are grateful to the Poul Due Jensen Foundation, which has supported the establishment of a new Centre for Digital Twin Technology at Aarhus University. Finally, we thank the reviewers for the thorough feedback.}

### 2020-11-13 background.tex
Updates from Overleaf

\subsection{FMU definitions}
To describe the formalization of FMUs, we adopt the vocabulary from \cite{gomes_lucio_vangheluwe_2019}. The main definitions of relevance to this paper will be presented, but readers are referred to the original publications for more information. This paper is only concerned with the initialization-phase of a co-simulation, making time of an FMU irrelevant. The formalization from Gomes et al. is extended with new definitions regarding algebraic loops, and convergence of fixed point iteration.
\begin{definition}[FMU]\label{def:fmu}
  An FMU with identifier $c$ is represented by the tuple   
  $$\tuple{\stateset{c}, \inputs{c}, \outputs{c}, \fset{c}, \fget{c}},$$
  where:
  \begin{inparadesc}
    \item $\stateset{c}$ represents the state space of FMU c;
    \item $\inputs{c}$ and $\outputs{c}$ the set of input and output variables, respectively;
    \item $\fset{c} : \stateset{c} \times \inputs{c} \times \values \to \stateset{c}$ and $\fget{c}: \stateset{c} \times \outputs{c} \to \values$ are functions to set the inputs and get the outputs, respectively (we abstract the set of values that each input/output variable can take as $\values$).
  \end{inparadesc}
\end{definition}

\begin{definition}[Scenario]\label{def:cosim_scenario}
  A scenario is a structure $\tuple{\fmus, \coupling}$ where each identifier $c \in \fmus$ is associated with an FMU, as defined in \cref{def:fmu}, and $\coupling(u)=y$ means that the output $y$ is connected to input $u$.
  Let $\allinputs = \bigcup_{c \in \fmus} \inputs{c}$ and $\alloutputs = \bigcup_{c \in \fmus} \outputs{c}$, then $\coupling : \allinputs \to \alloutputs$.
\end{definition}
Note it is possible for a single output to connect to a multiple inputs, but a single input can only rely on a single output.
%In Defn. 2, it seems that L is a function U --> Y, does that mean that each input variable feeds-through only one output variable?
The following definitions correspond to the operations that are permitted in the initialization phase of a co-simulation.

@@

\begin{definition}[Input Computation]\label{def:setin}
The $\fset{c}(\dontcare, \inputvar{c}, \inputV)$ represents the setting of input $\inputvar{c}$  of $c \in \fmus$. Given a co-simulation state, it checks whether all outputs connected to $\inputvar{c}$ are defined.
\end{definition}
%A faulty check will incur TODO

\begin{definition}[Fixed Point]\label{def:fixedpoint}
The $\fixedpoint{l}$ represents an order sequence of the setting or getting of all input and output variables of a given SCC $l$, see \cref{def:loop} for a definition of SCC.
The $\fixedpoint{l} \subseteq \bigcup_{c \in \fmus} \set{\fget{c},\fset{c}}$ 
\end{definition}

\begin{definition}[Initialization]\label{def:initialization}
  Given a scenario $\tuple{\fmus, \coupling}$, we define the initialization procedure $\sequence{\initcall_i}_{i \in \setnat}$ as is a finite ordered sequence of FMU function calls that needs to be performed in the initialization of a co-simulation scenario. The ordered sequence is defined as: $\sequence{\functioncall_i}_{i \in \setnat} = \functioncall_0, \functioncall_1, \ldots$ with
  $\functioncall_i \in \initcall_{} = \bigcup_{l \in loops} \fixedpoint{l},$
  and $i$ denoting the order of the function call.
  loops in defined as the set of all SCC see \cref{def:loop}.
\end{definition}

@@

\begin{figure}
    \centering
    \includegraphics[width=1\textwidth]{images/quarter_car_SCC.pdf}
    \caption{An FMU co-simulation scenario of the Quarter car and its Initialization Graph denoted with SCCs}
    \label{fig:fmu_cycle}
\end{figure}

@@

\begin{definition}[Algebraic loops] \label{def:loop}
An algebraic loop is defined as a non-trivial, strongly connected component of the graph in \cref{def:initialization_graph}.
Formally, a strong connected component satisfies $\set{a,b \in SCC: \mathit{Path}(a,b)}$, where $\mathit{Path}(a,b)$ is true when there's a path (including an empty path) between nodes $a$ and $b$ ($\mathit{Path}(a,a)$ is always true).
An $SCC$ is non-trivial when it has more than one node.
\end{definition}

Since the edges of the graph represent dependencies between variables, the value of every variable in a non-trivial strong component depends on itself.
Let $X$ denote a vector of one or more variables whose value depends on itself. The non-trivial strong component forms an equation with the form $F(X, U) = X$, where $F$ denotes the relations between the variables in the loop and $U$ denotes the variables whose values are calculated elsewhere. 
This means that algebraic loops need to be handled using fixed point iterations\cite{Gomes2018}.

@@

\subsection{INTO-CPS Maestro 2}
INTO-CPS Maestro 2\footnote{currently in alpha \url{https://github.com/INTO-CPS-Association/maestro/tree/2.0.0-alpha}}~\cite{Thule2020a} is a framework for creating simulation specifications and executing such specifications. The framework is FMI-based and set to supersede Maestro~\cite{Maestro} with the main goal of supporting research into co-simulation based on FMI.

@@

A specification is expressed in the domain-specific language called Maestro Base Language (MaBL), and it is explicit, such that the application of, i.e., FMUs are transparent. Expansion plugins can assist in creating such MaBL specifications, and one can apply expansion plugins that, in turn, generate the MaBL code. The plugin described in this paper is such an expansion plugin. The application of a plugin is evident in a MaBL specification. Upon processing of the specification, a new specification is created where the application of a plugin is replaced by the MaBL code generated by the plugin. This process is known as expansion, and a specification without any expansions remaining is a fully expanded MaBL specification.

To conduct a co-simulation, Maestro2 also features an interpreter that can execute a fully expanded MaBL specification, resulting in the execution of the co-simulation.

>>>>

\subsection{FMU Definitions}
To describe the formalization of FMUs, we adopt the vocabulary from \cite{gomes_lucio_vangheluwe_2019}. The main definitions of relevance to this paper will be presented, but readers are referred to the original publications for more information. This paper is only concerned with the initialization-phase of a co-simulation, making time of an FMU irrelevant. The formalization from Gomes et al.\cite{gomes_lucio_vangheluwe_2019} is extended with new definitions regarding algebraic loops, and convergence of fixed point iteration.
\begin{definition}[FMU]\label{def:fmu}
  An FMU with identifier $c$ is represented by the tuple   
  $$\tuple{\stateset{c}, \inputs{c}, \outputs{c}, \fset{c}, \fget{c}},$$
  where:
  \begin{inparadesc}
    \item $\stateset{c}$ represents the state space of FMU c;
    \item $\inputs{c}$ and $\outputs{c}$ the set of input and output variables, respectively;
    \item $\fset{c} : \stateset{c} \times \inputs{c} \times \values \to \stateset{c}$ and $\fget{c}: \stateset{c} \times \outputs{c} \to \values$ are functions to set the inputs and get the outputs, respectively (we abstract the set of values that each input/output variable can take as $\values$).
  \end{inparadesc}
\end{definition}

\begin{definition}[Scenario]\label{def:cosim_scenario}
  A scenario is a structure $\tuple{\fmus, \coupling}$ where each identifier $c \in \fmus$ is associated with an FMU, as defined in \cref{def:fmu}, and $\coupling(u)=y$ means that the output $y$ is connected to input $u$.
  Let $\allinputs = \bigcup_{c \in \fmus} \inputs{c}$ and $\alloutputs = \bigcup_{c \in \fmus} \outputs{c}$, then $\coupling : \allinputs \to \alloutputs$.
\end{definition}
Note a single output can connect to multiple inputs, but a single input can only rely on a single output.
The following definitions correspond to the operations that are permitted in the initialization phase of a co-simulation.

@@

\begin{definition}[Input Computation]\label{def:setin}
The $\fset{c}(\dontcare, \inputvar{c}, \inputV)$ represents the setting of input $\inputvar{c}$  of $c \in \fmus$. Given a co-simulation state, it checks whether all outputs connected to $\inputvar{c}$ are defined.
\end{definition}

\begin{definition}[Fixed Point]\label{def:fixedpoint}
The $\fixedpoint{l}$ represents an ordered sequence of the setting or getting of all variables of a given SCC $l$, see \cref{def:loop} for a definition of SCC.
The $\fixedpoint{l} \subseteq \bigcup_{c \in \fmus} \set{\fget{c},\fset{c}}$ 
\end{definition}

\begin{definition}[Initialization]\label{def:initialization}
  Given a scenario $\tuple{\fmus, \coupling}$, we define the initialization procedure $\sequence{\initcall_i}_{i \in \setnat}$ as is a finite ordered sequence of FMU function calls that needs to be performed in the initialization of a co-simulation scenario. The ordered sequence is defined as: $\sequence{\functioncall_i}_{i \in \setnat} = \functioncall_0, \functioncall_1, \ldots$ with
  $\functioncall_i \in \initcall_{} = \bigcup_{l \in loops} \fixedpoint{l},$
  and $i$ denoting the order of the function call.
  Loops is defined as the set of all SCC see \cref{def:loop}.
\end{definition}

@@

\begin{figure}
    \centering
    \includegraphics[width=1\textwidth]{images/quarter_car_SCC.pdf}
    \caption{An FMU co-simulation scenario of the Quarter car and its Initialization Graph denoted with SCCs.}
    \label{fig:fmu_cycle}
\end{figure}

@@

\begin{definition}[Algebraic loops] \label{def:loop}
An algebraic loop is defined as a non-trivial, strongly connected component of the graph in \cref{def:initialization_graph}.
Formally, a strong connected component satisfies $\set{a,b \in SCC: \mathit{Path}(a,b)}$, where $\mathit{Path}(a,b)$ is true when there's a path (including an empty path from a node to itself) between nodes $a$ and $b$ ($\mathit{Path}(a,a)$ is always true).
An $SCC$ is non-trivial when it has more than one node.
\end{definition}

Since the edges of the graph represent dependencies between the variables, the value of every variable in a non-trivial strong component depends on itself.
Let $X$ denote a vector of one or more variables whose value depends on itself. The non-trivial strong component forms an equation with the form $F(X, U) = X$, where $F$ denotes the relations between the variables in the loop and $U$ denotes the variables whose values are calculated elsewhere. 
This means that algebraic loops need to be handled using fixed point iterations\cite{Gomes2018}.

@@

\subsection{INTO-CPS Maestro 2}
INTO-CPS Maestro 2\footnote{currently in alpha \url{https://github.com/INTO-CPS-Association/maestro/tree/2.0.0-alpha}.}~\cite{Thule2019b} is a framework for creating simulation specifications and executing such specifications. The framework is FMI-based and set to supersede Maestro~\cite{Maestro} with the main goal of supporting research into co-simulation based on FMI.

@@

A specification is expressed in the domain-specific language called Maestro Base Language (MaBL), and it is explicit, such that the application of, i.e., FMUs are transparent. Expansion plugins can assist in creating such MaBL specifications, and one can apply expansion plugins that, in turn, generate the MaBL code. The plugin described in this paper is such an expansion plugin. The application of a plugin is evident in a MaBL specification. Upon processing of the specification, a new specification is created where the application of a plugin is replaced by the MaBL code generated by the plugin. This process is known as expansion, and a specification without any expansions remaining is a fully expanded MaBL specification. An example of a part of the folded MaBL specification of the case study example of \cref{sec:case_study} can be seen below.

\begin{lstlisting}[language=C++]
simulation
import Initializer;
{
FMI2 chassis = load("FMI2", "{8c4e810f-3df3-4a00-8276-176fa3c9f000}", "src/chassis-c.fmu");
...
IFmuComponent components[3]={chassis,suspension, road};
expand initialize(components,START_TIME, END_TIME);
...
}
\end{lstlisting}
To conduct a co-simulation, Maestro2 also features an interpreter that can execute a fully expanded MaBL specification, resulting in the execution of the co-simulation.

### 2020-11-13 conclusion.tex
Updates from Overleaf

\section{Concluding remarks}\label{sc:summary}
This work uses a topological ordering of a dependency graph of the interconnected FMUs variable and internal FMU connections along with predicates from the FMI specification to calculate a correct initialization order for a co-simulation scenario potentially containing algebraic loops.
The initialization procedure optimizes the initialization order by grouping variables with similar characteristics to perform the fewest possible operations in the initialization procedure.
This approach supports the initialization of a co-simulation scenario containing algebraic loops by using fixed point iteration. 
%The approach discards all co-simulation scenarios, not adhering to the law of convergence defined in \cref{def:convergence}.
The approach is suitable to combine with well-established master algorithms like Gauss-Seidel and Jacobi \cite{Palensky2017}. 
The approach is realized as an expansion plugin for the open-source INTO-CPS Maestro 2 tool and verified against the existing \textit{Initializer} and the calculated initialization order was verified against an established co-simulation Algorithm Generator and Verifier implemented in Prolog\cite{gomes_lucio_vangheluwe_2019}\\
Future work includes formal verification of the plugin using the Logika framework\cite{inbook}.
We will also look into the generation of a verification strategy for the whole Maestro 2 framework to examine how different forms of verification jointly can extend the trust of the correctness of the result of a co-simulation.

\paragraph*{\textbf{Acknowlegements.}}We would like to thank Stefan Hallerstede, Christian Møldrup Legaard, and Peter Gorm Larsen for providing valuable input to this paper and the developed plugin.

>>>>

\section{Concluding Remarks}\label{sc:summary}
This work uses a topological ordering of a dependency graph of the interconnected FMUs variable and internal FMU connections along with predicates from the FMI specification to calculate a correct initialization order for a co-simulation scenario potentially containing algebraic loops.
The initialization procedure optimizes the initialization order by grouping variables with similar characteristics to perform the fewest possible operations in the initialization procedure.
This approach supports the initialization of a co-simulation scenario containing algebraic loops by using fixed point iteration. 
%The approach discards all co-simulation scenarios, not adhering to the law of convergence defined in \cref{def:convergence}.
The approach is suitable to combine with well-established master algorithms like Gauss-Seidel and Jacobi \cite{Palensky2017}. 
The approach is realized as an expansion plugin for the open-source INTO-CPS Maestro 2 tool and verified against the existing \textit{Initializer} and the calculated initialization order was verified against an established co-simulation Algorithm Generator and Verifier implemented in Prolog\cite{gomes_lucio_vangheluwe_2019}\\
Future work includes formal verification of the plugin using the Logika framework\cite{inbook}.
We will also look into the generation of a verification strategy for the whole Maestro 2 framework to examine how different forms of verification jointly can extend the trust of the correctness of the result of a co-simulation.

\subsubsection*{Acknowledgements} We would like to thank Stefan Hallerstede, Christian Møldrup Legaard, and Peter Gorm Larsen for providing valuable input to this paper and the developed plugin.
